\thesistitle

### 2019-08-21 sections/abstract.tex
Introduction

\chapter*{Abstract}
%\todo[inline,color=blue!10]{This should be done when most of the rest of the document is finished. Be concise, introduce context, problem, known approaches, your solution, your findings.}
Duplication in source code is often seen as one of the most harmful types of technical debt as it increases the size of the codebase and creates implicit dependencies between fragments of code. To remove such anti-patterns, the codebase should be refactored. Many tools aid in the detection process of such duplication problems, but fail to determine whether refactoring a duplicate fragment would improve the maintainability.

>>>>

\chapter*{Abstract}
%\todo[inline,color=blue!10]{This should be done when most of the rest of the document is finished. Be concise, introduce context, problem, known approaches, your solution, your findings.}
Duplication in source code is often seen as one of the most harmful types of technical debt as it increases the size of the codebase and creates implicit dependencies between fragments of code. To remove such anti-patterns, the codebase should be refactored. Many tools aid in the detection process of such duplication problems but fail to determine whether refactoring a duplicate fragment would improve the maintainability.

### 2019-08-21 sections/introduction.tex
Introduction

\chapter{Introduction}
%\todo[inline,color=blue!10]{Context: what is the bigger scope of the problem you are trying to solve? Try to connect to societal/economical challenges.
%Problem Analysis: Here you present your analysis of the problem situation that your research will address.
%How does this problem manifest itself at your host organisation?
%Also summarises existing scientific insight into the problem.}
\label{ch:introduction}
% TODO: Sander - What is refactoring? Hier zeg je meteen waarvoor het gebruikt wordt
Refactoring is the process of restructuring code to improve quality related attributes of a codebase (maintainability, performance, etc.) without changing the functionality. There are many methods that have been introduced to help with the process of refactoring \cite{fowler2018refactoring, wake2004refactoring}. However, most of these methods still require manual assessment of where and when to apply them. Because of this, refactoring takes up a signification portion of the development process \cite{lientz1978characteristics, mens2004survey}, or does not happen at all \cite{mens2003refactoring}. %For a large part, refactoring requires domain knowledge to do it right. However, there are also refactoring opportunities that are rather trivial and repetitive to execute.
% De zien hierna staat ook al in laatste alineas, past minder goed binnen intro imo
%In this thesis, we take a look at the challenges and opportunities in automatically refactoring duplicated code, also known as ``code clones''. The main goal is to improve maintainability of the refactored code.

@@

@@

\subsection{The problem}
The maintainability of a codebase has a large impact on the time and effort spent on building the desired software system \cite{bakota2012cost, munson1978software}. The maintainability of software is one of the factors to be kept under control in order to avoid major delays and unexpected costs as a result of a software project \cite{fowler2018refactoring}. One factor that has a major impact on the maintainability of a software system is the amount of duplicate code present in a codebase \cite{heitlager2007practical, fowler1999refactoring}.

The process of improving maintainability through the refactoring of duplicate code is time consuming and error-prone. This process mainly consists of these two aspects:
\begin{itemize}
	\item Find refactoring candidates, either tool-assisted or fully manually. % TODO: Sander - is tool assisted niet ook beetje manual? lijkt meer tool-assisted vs fully manual of iets
	\item Refactor identified candidates, either tool-assisted or fully manually.
\end{itemize}
There are tools that assist in the process of finding duplicate code, but none of these tools identify all duplication problems. Often, these clone detection tools result in false-positives and false negatives \cite{roy2007survey}. We define false negatives as clones that can and should be refactored but are not found due to the configuration by which they are detected. We define false-positives as clones that do not improve the system maintainability when refactored. Many false negatives are due to the thresholds that are used. Many clone detection tools \cite{sajnani2016sourcerercc, svajlenko2016bigcloneeval} only consider clones that have a minimum of 6 lines. Using such an approach is bound to result in many false negatives, but using a lower threshold is shown to increase the number of false-positives. %TODO cite threshold paper

A study by Batova et al. \cite{bavota2012does} shows that the process of refactoring often leaves side effects in the code. This study reports that refactorings involving hierarchies (e.g., pull up method) % TODO: Sander -unsure of pull up method bekende term is, miss "creating a new method in a superclass"
, tend to induce faults very frequently. \textit{This suggests more accurate code inspection or testing activities % TODO: Sander - .. are needed?
when such specific refactorings are performed.} Such specific refactorings often have to be used when dealing with duplication in software \cite{fowler2018refactoring, fontana2015duplicated}. Because of that, refactoring code clones has been empirically proven to cause bugs or other side effects in code.

@@

\section{Scope}
% TODO: Sander - iets over OOP hier?
In this study, we perform research efforts to be able to detect refactorable clones. We will apply refactoring techniques to a subset of these clones and analyze the maintainability of the resulting source code.

There is a lot of study on how what metrics to consider to measure maintainability. In this study, we focus solely on the practical maintainability model by Heitlager et al. \cite{heitlager2007practical}. We consider the maintainability scores described in that paper as an sufficiently accurate indication of maintainability. This will be used to quantitatively determine the impact on maintainability when applying refactoring techniques to code clones.

Applying refactoring often entails creating new method declarations, class declarations, etc. Each of these declarations needs to have a name. Because the quality of the name is not included in the maintainability metrics we use, finding appropriate names for automatically refactored code fragments is out of the scope for this study. For our automated refactoring efforts, we will use generated names for these declarations.

It is very disputable whether unit tests apply to the same maintainability metrics that applies to the functional code. Because of that, for this research, unit tests are not taken into scope. The findings of this research may be applicable to those classes, but we will not argue the validity.

\section{Outline}
In Chapter~\ref{ch:background} we describe the background of this thesis. In Chapter~\ref{chap:clonetypes} we list shortcomings with clone definitions from literature and propose a set of clone type definitions which can be automatically refactored. % TODO: Sander - Nu zeg je dat de definities automatisch gerefactored kunnen worden
In Chapter \ref{ch:clonerefactor} we propose a tool to detect, analyze and automatically refactor such clones that can be refactored. Using this tool we perform a set of experiments, of which the results are shown in Chapter~\ref{ch:results} and discussed in Chapter~\ref{ch:discussion}. Chapter~\ref{ch:related_work} contains the work related to this thesis. Finally, we present our concluding remarks in Chapter~\ref{ch:conclusion} together with future work.

>>>>

\chapter{Introduction}
%\todo[inline,color=blue!10]{Context: what is the bigger scope of the problem you are trying to solve? Try to connect to societal/economical challenges.
%Problem Analysis: Here you present your analysis of the problem situation that your research will address.
%How does this problem manifest itself at your host organisation?
%Also summarises existing scientific insight into the problem.}
\label{ch:introduction}
Refactoring is the process of restructuring code to improve quality related attributes of a codebase (maintainability, performance, etc.) without changing the functionality. There are many methods that have been introduced to help with the process of refactoring \cite{fowler2018refactoring, wake2004refactoring}. However, most of these methods still require manual assessment of where and when to apply them. Because of this, refactoring takes up a signification portion of the development process \cite{lientz1978characteristics, mens2004survey}, or does not happen at all \cite{mens2003refactoring}. %For a large part, refactoring requires domain knowledge to do it right. However, there are also refactoring opportunities that are rather trivial and repetitive to execute. In this thesis, we take a look at the challenges and opportunities in automatically refactoring duplicated code, also known as ``code clones''. The main goal is to improve maintainability of the refactored code.

@@

\todo{A paragraph aboout current clone types being ill-defined for automated refactoring}

@@

\subsection{The problem}
The maintainability of a codebase has a large impact on the time and effort spent on building the desired software system \cite{bakota2012cost, munson1978software}. The maintainability of software is one of the factors to be kept under control in order to avoid major delays and unexpected costs as the outcome of a software project \cite{fowler2018refactoring}. One factor that has a major impact on the maintainability of a software system is the amount of duplicate code present in a codebase \cite{heitlager2007practical, fowler1999refactoring}.

The process of improving maintainability through the refactoring of duplicate code is time consuming and error-prone. This process mainly consists of two aspects:
\begin{itemize}
	\item Find refactoring candidates, either tool-assisted or fully manually.
	\item Refactor identified candidates, either tool-assisted or fully manually.
\end{itemize}
There are tools that assist in the process of finding duplicate code, but none of these tools identify all duplication problems. Often, these clone detection tools result in false-positives and false negatives \cite{roy2007survey}. We define false negatives as clones that can and should be refactored but are not found due to the configuration by which they are detected. We define false-positives as clones that do not improve the system maintainability when refactored. Many false negatives are due to the thresholds that are used. Clone detection tools are configured using a predefined amount of lines or tokens that two duplicate fragments should minimally have for them to be considered clones \cite{sajnani2016sourcerercc, svajlenko2016bigcloneeval}. Often, such thresholds are intuitively chosen \cite{li2006cp, roy2009mutation}, leading to a potentially non-optimal selection of clones that are presented to developer.

A study by Batova et al. \cite{bavota2012does} shows that the process of refactoring often leaves side effects in the code. This study reports that refactorings moving components in the inheritance hierarchy of object-oriented software tend to induce faults very frequently. This suggests that more accurate code inspection or testing activities are needed when such refactoring techniques are performed. Those refactoring techniques often have to be used when dealing with duplication in software \cite{fowler2018refactoring, fontana2015duplicated}. Because of that, refactoring code clones has been empirically proven to cause bugs or other side effects in code.

@@

\section{Scope}
In this study, we perform research efforts to be able to detect refactorable clones. We will apply refactoring techniques to a subset of these clones and analyze the maintainability of the resulting source code.

\subsection{Java}
This research is limited to object-oriented programming languages. We relate our research to the most popular object-oriented programming languages (Python, Java, C\#). Our experiments and results have been conducted in the Java programming language, because many refactoring practises are language-specific.

\subsection{Measuring Maintainability}
There is a lot of study on how what metrics to consider to measure maintainability. In this study, we focus solely on the practical maintainability model by Heitlager et al. \cite{heitlager2007practical}. We consider the maintainability scores described in that paper as an sufficiently accurate indication of maintainability. This will be used to quantitatively determine the impact on maintainability when applying refactoring techniques to code clones.

\subsection{Naming Declarations}
By our automated refactoring model new method declarations, class declarations and interfaces are created. Each of these declarations needs to have a name. Because the quality of the name will not have an impact on the maintainability metrics we use, finding appropriate names for automatically refactored code fragments is out of the scope for this study. For our automated refactoring efforts, we will use generated names for these declarations. %This may decrease the usefulness of the results of this research for refactoring-assistance.

\subsection{Testcode}
It is very disputable whether unit tests apply to the same maintainability metrics that applies to the functional code. Because of that, for this research, unit tests are not taken into scope. The findings of this research may be applicable to those classes, but we will not argue the validity. Furthermore, any other code than the production code of the system will not be taken into account. In Section~\ref{chap:corpus} we describe what classes we perform our experiments on.

\section{Outline}
In Chapter~\ref{ch:background} we describe the background of this thesis. In Chapter~\ref{chap:clonetypes} we list shortcomings with established clone definitions which make them less suitable for automated refactoring. We also propose a set of clone type definitions that strive to solve these shortcomings. In Chapter \ref{ch:clonerefactor} we propose a tool to detect, analyze and automatically refactor such clones that can be refactored. Using this tool we perform a set of experiments, of which the setup is explained in Chapter~\ref{ch:experimentalsetup}. The results of the experiments are in shown in Chapter~\ref{ch:results} and discussed in Chapter~\ref{ch:discussion}. Chapter~\ref{ch:related_work} contains the work related to this thesis. Finally, we present our concluding remarks in Chapter~\ref{ch:conclusion} together with future work.
